Stack Safety as a Security Property

### 2020-06-27 main.tex
Stack safety section with explanation that actually flows and makes some sense - at least to me

The first part of stack safety is {\em local state encapsulation}.
Informally, it captures the notion that a caller's data is
protected from a callee's actions. We formalize this notion
as a security property with a confidentiality and an integrity
component, where the former protects the caller's data from being
read and the latter from being written.

As is standard with security properties, we associate with each
component of the machine a {\em label}: a pair of security levels
denoting high or low confidentiality and high or low integrity.
%
\[\mathit{label} ::= \{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT,\LOWINT\}\]
%
Since we need to enforce integrity and confidentiality at {\em every}
function call in a trace, we must treat the security level of data
dynamically. To that end, we introduce the notion of a {\em contour},
a map from components to labels:
%
\[\contour \in \CONTOURS ::= \COMPONENTS \rightarrow \mathit{label}\]

@@

@@

Confidentiality is a noninterference-style property, comparing the
behavior of a machine with an arbitrary variant that preserves low
confidentiality data.

\leo{Variations or variants?}
\definition{Variations}

We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
another given a contour $\contour$ if they agree on low
confidentiality data:
%
\[
\mach \approx_\contour \mach' \triangleq \forall \component .
\contour(\component) = \LOWSEC \rightarrow \mach(\component) =
\mach'(\component)
\]

Let $\callmap$ be a call map and $\codemap$ a code map.
Confidentiality with respect to $\callmap$ and $\codemap$ should be
enforced at {\em every} transition between a caller $f$ and a callee
$f'$.
%
Suppose that an execution trace starting from some initial
machine-policy state pair contains such a transition. That is, it
contains some step between state state pairs $(\mach_c,\pol_c)
\stepstoobs{\obs}_P (\mach_e, \pol_e)$ for some observation $\obs$
such that $\mach_c$ is a valid call with $n$ arguments
($\callmap(\mach_c(\PCname)) = n$) and $\mach_e$ is an entry point for
$f'$ ($\codemap(\mach_e(\PCname)) = f'$).

Now let $MP$ be the (potentially infinite) prefix of the trace
starting at the entry $(\mach_e, \pol_e)$ until the corresponding
return: $(\mach_e, \pol_e) \hookrightarrow MP ~ | ~ \ret{\mach_c}$ and
$\contour$ be the contour at the call: $\contour =
\mathit{Cof}(\mach_c, f', n)$.

@@

We will say that $MP$ satisfies {\em trace confidentiality} with
respect to $\callmap$, $\codemap$, and $\contour$ if, for every
$\contour$-variation $\mach_e'$ of the entry state, and the prefix $M$
of its induced traced until the corresponding return $\mach_e'
\hookrightarrow M ~ | ~ \ret{\mach_c}$, one of three cases holds:
\begin{itemize}
\item If the call successfully returns, that is $MP$ is finite and its
  final state $\mach_r$ is a valid return, then $M$ must also be
  finite with some final state $\mach_r'$ that has changed in the same
  manner:
$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
  \vee \mach_e'(\component) \not = \mach_r'(\component) \rightarrow
  \mach_r(\component) = \mach_r'(\component)$$

  In addittion their observations must be equivalent:
  $$\pi_o(MP) \eqsim \pi_o(M)$$
\item If the call trace never returns and is infinite, then the
  variant trace must also never return and their observations are
  equivalent:
  $$\pi_o(MP) \eqsim \pi_o(M)$$  
\item Finally, if the call trace ends prematurely with a policy fault,
  then we require that the observations of $MP$ are a prefix of those
  of the variation:
  $$\pi_o(MP) \lesssim \pi_o(M)$$
\end{itemize}

@@

We say that a system satisfies {\em stack confidentiality} with
respect to a call map $\callmap$ and a code map $\codemap$ if for
every initial machine-policy pair $(\mach_0, \pol_0)$ and its induced
trace, every call subtrace\leo{Can i reuse stuff above?} in it
satisfies trace confidentiality.

  A system satisfies eager stack integrity with respect to a call map
  \(\callmap\) and code map \(\codemap\) if for any initial mpstate
  \(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each
  adjacent pair of a call state \(\mach_c\) and state \(\mach_e\) with policy
  state \(\pol_e\) in the induced trace from \(\mpstatename\), where
  \(\callmap(\mach_c(\PCname)) = n\) and \(\codemap(\mach_e(\PCname)) =
  \funid\), if we take the prefix until a return
  \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_c}\) and
  \(\MPT\) ends with some \(\mach_p\), \(\mach_p\) matches \(\mach_c\) on each
  component whose integrity in \(\mathit{Cof}(\mach_c,\funid,n)\) is
  \(\HIGHINT\).

>>>>

The first part of stack safety is {\em local state encapsulation}.
Informally, it captures the notion that a caller's data is
protected from a callee's actions. We formalize this notion
as a security property with an integrity and a confidentiality
component, where the former protects the caller's data from being
written and the latter from being read.

As is standard with security properties, we associate with each
component of the machine a {\em label}: a pair of security levels
denoting high or low confidentiality and high or low integrity.
%
\[\mathit{label} ::= \{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT,\LOWINT\}\]
%
In a traditional setting, integrity would enforce that the contents of
$\HIGHINT$ locations remain unchanged, while confidentiality would
enforce that the contents of $\HIGHSEC$ locations don't influence the
observable behavior. However, when dealing with nested calls and
returns, stack safety needs to enforce integrity and confidentiality
for {\em every} function call in a trace; that is, we must treat the
security level of data dynamically.
%
To that end, we introduce the notion of a {\em contour}, a map from
components to labels:
%
\[\contour \in \CONTOURS ::= \COMPONENTS \rightarrow \mathit{label}\]

@@

\paragraph*{Stack Integrity}

The first component of local state encapsulation is an integrity
property: it ensures that the caller's data is never overwritten by a
callee. We will formally define integrity in two steps: first, we will
define integrity as a trace property over the trace corresponding to a
single call (from its entry point to its corresponding return), and
then we will lift that to a property for every such subtrace of a
callee.

\definition{Trace Integrity}
\label{def:trace-integrity}

Let $\contour$ be a contour and $MP$ be a (potentially infinite)
$\MPOTRACE$. We will say $MP$ satisfies {\em trace integrity} with
respect to $\contour$ if its final state (if any) agrees with its
initial one on all components marked $\HIGHSEC$ in $\contour$:
%
\[
\mathit{int} ~ \contour ~ MP \triangleq
\left \{ \begin{array}{ll}
  \mathit{True}, & \text{if $MP$ is infinite}\\
  \forall \component. ~ \contour(\component) = \HIGHSEC \rightarrow m_0(\component) = m_f(\component), & \text{otherwise} \\
\end{array}
\right .
\]  
\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}

Intuitively, the caller expects to find their private data untouched
after the callee returns. That means that if the callee never returns,
integrity is trivially satisfied. Similarly, the callee is allowed to
temporarily overwrite these locations, as long as they are restored
once control returns. We will see variations of this later on: a
stronger, inductive integrity property that ensures that the caller's
private data is never overwritten, even during the caller's execution
(a variation that leads to easier reasoning and more effective testing);
and a weaker, lazy integrity property that only ensures that the
locations that a caller accesses are preserved (a variation that allows
for a more efficient enforcement mechanism).

Trace integrity characterizes the trace corresponding to a single
callee. We now lift this definition to an integrity property for a
complete execution trace. We do this by requiring that trace integrity
holds for every transition between a caller and a callee $f$ for
the entirety of the trace of $f$.

\definition{Transition}

Given a trace $MP$, a call map $\callmap$ and a code map $\codemap$,
an $n-${\em transition} to a callee $f$ is a pair of machine-policy
state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ in the trace
such that:
\begin{itemize}
\item one state pair steps to the other: 
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstoobs{\obs}_P (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
system satisfies {\em stack integrity} with respect to $\callmap$ and
$\codemap$ if, for every initial machine-policy state pair $(\mach_0,
\pol_0)$ and its induced trace $MP$, and every $n$-transition
$(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee $f$ in $MP$,
the prefix $MP_{pre}$ of $MP$ from $(\mach_e, \pol_e)$ until the first
return corresponding to $\mach_c$ (if any) satisfies trace integrity
with respect to the contour at the call:
%
$$\forall MP_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
MP_{pre} ~ | ~ \ret{\mach_c} \rightarrow
\mathit{int}~\mathit{Cof}(\mach_c, f, n)~MP_{pre}$$

@@

The second component of local state encapsulation is a confidentiality
property: just like integrity ensures that a caller's data is never
overwritten by a callee, confidentiality ensures that it never
influences its observable behavior. We will formally define
confidentiality in two similar steps: a trace version and a system
version.

Unlike integrity, confidentiality is a noninterference-style property,
comparing the behavior of a machine with an arbitrary variant that
preserves low confidentiality data.

\definition{Variations}

We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
another given a contour $\contour$ if they agree on low
confidentiality data:
%
\[
\mach \approx_\contour \mach' \triangleq \forall \component .
\contour(\component) = \LOWSEC \rightarrow \mach(\component) =
\mach'(\component)
\]
%
Then, the trace version of confidentiality operates on two traces: the
``original'' one $MP \in \MPOTRACE$ that corresponds to a callee $f$
and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
if it started from a variant initial machine state.

@@

\leo{Add info for ``valid return'':}
We will say two traces $MP$ and $M$ satisfy {\em trace
  confidentiality} with respect to $\contour$ if:
\begin{itemize}
\item If $MP$ is finite and its final state $\mach_r$ is a valid
  return, then $M$ is also finite with some final state $\mach_r'$ and
  any changes from the initial states have changed in the same way:
$$\forall \component. ~ \mach_e(\component) \not = \mach_r(\component)
  \vee \mach_e'(\component) \not = \mach_r'(\component) \rightarrow
  \mach_r(\component) = \mach_r'(\component)$$
  In addittion their observations must be equivalent:
  $$\pi_o(MP) \eqsim \pi_o(M)$$
\item If $MP$ is infinite, then $M$ is also infinite and their
  observations are equivalent:
  $$\pi_o(MP) \eqsim \pi_o(M)$$  
\item If the $MP$ ends prematurely with a policy fault, then
  we require that its observations are a prefix of those of $M$:
  $$\pi_o(MP) \lesssim \pi_o(M)$$
\end{itemize}

\leo{Discuss variations in a similar manner}

To lift trace confidentiality to a system property, we follow the same
pattern as in stack integrity:

@@

Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
system satisfies {\em stack confidentiality} with respect to
$\callmap$ and $\codemap$ if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $MP$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
$f$ in $MP$, the prefix $MP_{pre}$ of $MP$ from $(\mach_e, \pol_e)$
until the first return corresponding to $\mach_c$ (if any) satisfies
trace integrity with respect to every similarly prefixed machine trace
of every variation $\mach_e'$ with the contour at the call:
$$\forall MP_{pre} ~ M_{pre}. ~
\mpstate{\mach_e}{\pol_e} \hookrightarrow_P MP_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
\mach_e' \hookrightarrow M_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ MP_{pre} ~ M_{pre}$$
